A Survey of Real-Time Strategy Game AI\\ Research and Competition in StarCraft

### 2013-03-11 survey.tex
A few updates in the conclusions

\begin{figure*}[tb]
    \centering
    \includegraphics[width=\textwidth]{figures/botLadder1.pdf}
    \caption{StarCraft Bot Ladder: bot's ELO Rating from February 1, 2012 to June 20, 2012}
    \label{fig:botLadder1}
\end{figure*}

\begin{figure*}[tb]
    \centering
    \includegraphics[width=\textwidth]{figures/botLadder2.pdf}
    \caption{StarCraft Bot Ladder: bot's ELO Rating from September 1, 2012 to January 10, 2013}
    \label{fig:botLadder2}
\end{figure*}

@@

The StarCraft Bot Ladder is a website\footnote{\url{http://bots-stats.krasi0.com}} where bot versus bot matches are automatized, and are running all the time. This ladder is a great resource for creating data sets (all the game replays are available) and statistics. For bot ranking, the ladder uses an ELO rating system suitable for calculating the \emph{relative skill level} of a bot in two-player games. In the ELO system each player has a numerical rating that gets incremented or decremented some points after each game. The amount of points depends on the difference in the ratings of the players. A player will gain more points by beating a higher-rated player than by beating a lower-rated player. These kind of rating system is widely used in games like chess. This bot ladder compiles different versions of bots from the main worldwide competitions (like AIIDE, CIG or SSCAI\footnote{\url{http://sscaitournament.com}}), even some independent or ``under construction'' bots. Therefore, it is a very good resource to test the performance of new bots against the current state of the art in StarCraft bots before participating in the official competitions.

Figure \ref{fig:botLadder1} shows the ELO rating of the bots in the ladder during the first half year of 2012. The ranking is practically equal than the AIIDE 2011 competition, showing the lack of adaptability of the current bots. We can notice these more extremely in Figure \ref{fig:botLadder2} for the second half year of 2012. During this period new bots were introduced in the ladder. We can observe how the first version of KillerBot made a huge impact on Skynet ranking and finally, the second version of KillerBot quickly became the best bot for more than one month (again we can see how the rest of the bots aren't able to adapt and the ranking doesn't change so much). And finally, in January, the Ximp bot appears with a new strategy that overcomes the rest of the bots. Both KillerBot and Ximp use hard-coded strategies without any kind of adaptation capabilities. However, they implement strategies that no other bot has a counter for, and thus manage to win a very large percentage of games. This points out, once again, that one of the major open challenges in RTS game AI is how achieving adaptive strategies, that can recognize the opponent's intentions, and select an adequate response.

@@

As the list in the previous section indicates, Real-Time Strategy games are an excellent testbed for AI techniques, which pose a very large list of open problems. As Section \ref{sec:competition} has shown, the current top performing programs to play RTS games such as StarCraft still rely mainly on hard-coded strategies. It is still possible to perform strongly, or even win, one of these competitions simply by finding a hard-coded strategy that no other bot has a predefined counter-strategy for. Additionally, good human players are still clearly superior to the best computer programs. Compared to other genres of game AI, this is one of the few cases where bots simply have to improve to become more believable and thus more fun to play against for human players.

>>>>

\begin{figure*}[tb]
    \centering
    \includegraphics[width=\textwidth]{figures/botLadder1.pdf}
    \caption{StarCraft Bot Ladder: bot's Elo Rating from February 1, 2012 to June 20, 2012}
    \label{fig:botLadder1}
\end{figure*}

\begin{figure*}[tb]
    \centering
    \includegraphics[width=\textwidth]{figures/botLadder2.pdf}
    \caption{StarCraft Bot Ladder: bot's Elo Rating from September 1, 2012 to January 10, 2013}
    \label{fig:botLadder2}
\end{figure*}

@@

The StarCraft Bot Ladder is a website\footnote{\url{http://bots-stats.krasi0.com}} where bot versus bot matches are automatized, and are running all the time. This ladder is a great resource for creating data sets (all the game replays are available) and statistics. For bot ranking, the ladder uses an Elo rating system suitable for calculating the \emph{relative skill level} of a bot in two-player games. In the Elo system each player has a numerical rating that gets incremented or decremented some points after each game. The amount of points depends on the difference in the ratings of the players. A player will gain more points by beating a higher-rated player than by beating a lower-rated player. These kind of rating system is widely used in games like chess. This bot ladder compiles different versions of bots from the main worldwide competitions (like AIIDE, CIG or SSCAI\footnote{\url{http://sscaitournament.com}}), even some independent or ``under construction'' bots. Therefore, it is a very good resource to test the performance of new bots against the current state of the art in StarCraft bots before participating in the official competitions.

Figure \ref{fig:botLadder1} shows the Elo rating of the bots in the ladder during the first half year of 2012. The ranking is practically equal than the AIIDE 2011 competition, showing the lack of adaptability of the current bots. We can notice these more extremely in Figure \ref{fig:botLadder2} for the second half year of 2012. During this period new bots were introduced in the ladder. We can observe how the first version of KillerBot made a huge impact on Skynet ranking and finally, the second version of KillerBot quickly became the best bot for more than one month (again we can see how the rest of the bots aren't able to adapt and the ranking doesn't change so much). And finally, in January, the Ximp bot appears with a new strategy that overcomes the rest of the bots. Both KillerBot and Ximp use hard-coded strategies without any kind of adaptation capabilities. However, they implement strategies that no other bot has a counter for, and thus manage to win a very large percentage of games. This points out, once again, that one of the major open challenges in RTS game AI is how achieving adaptive strategies, that can recognize the opponent's intentions, and select an adequate response.

@@

As the list in the previous section indicates, Real-Time Strategy games are an excellent testbed for AI techniques, which pose a very large list of open problems. As Section \ref{sec:competition} has shown, the current top performing programs to play RTS games such as StarCraft still rely mainly on hard-coded strategies. It is still possible to perform strongly, or even win, one of these competitions simply by finding a hard-coded strategy that no other bot has a predefined counter-strategy for. Additionally, good human players are still clearly superior to the best computer programs. From an industry point of view, one additional challenge is to make bots more believable to play against, and thus, more fun for human players (this includes, for example, doing actual scouting, instead of cheating and having full information of the game state).
